other functionalities

### 2022-01-10 inst/gitbook/05_other-functionalities.Rmd
added match.tip.edge entry

>>>>

## `match.tip.edge`

This function matches a vector of discreet tip values with the edges connecting these tips in the `"phylo"` structure.
This can be used to pull the branches of interest for some specific trait of some group of species or for colouring tree tips based on clades.

For example, with the `charadriiformes` dataset, you can plot the tree with the branches coloured by clade.
To work properly, the function requires the characteristics of the tip labels (e.g. the clade colour) to match the order of the tips in the tree:

```{r}
## Loading the charadriiformes data
data(charadriiformes) 
## Extracting the tree
my_tree <- charadriiformes$tree
## Extracting the data column that contains the clade assignments
my_data <- charadriiformes$data[, "clade"]
## Changing the levels names (the clade names) to colours
levels(my_data) <- c("orange", "blue", "darkgreen")
my_data <- as.character(my_data)
## Matching the data rownames to the tip order in the tree
my_data <- my_data[match(ladderize(my_tree)$tip.label, rownames(charadriiformes$data))]
```

We can then match this tip data to their common descending edges.
We will also colour the edges that is not descendant directly from a common coloured tip in grey using `"replace.na = "grey"`.
Note that these edges are usually the edges at the root of the tree that are the descendant edges from multiple clades.

```{r}
## Matching the tip colours (labels) to their descending edges in the tree
## (and making the non-match edges grey)
clade_edges <- match.tip.edge(my_data, my_tree, replace.na = "grey")

## Plotting the results
plot(ladderize(my_tree), show.tip.label = FALSE, edge.color = clade_edges)
```

But you can also use this option to only select some specific edges and modify them (for example making them all equal to one):

```{r}
## Adding a fixed edge length to the green clade
my_tree_modif <- my_tree
green_clade <- which(clade_edges == "darkgreen")
my_tree_modif$edge.length[green_clade] <- 1
plot(ladderize(my_tree_modif), show.tip.label = FALSE, edge.color = clade_edges)
```
